### 2021-11-24 draft.tex
Deploying to gh-pages from @ gottacatchenall/thesis_proposal@043949f50c1aaf922b2452842270c69e5ba20c29 🚀

The second major challenge in ecological forecasting is that the
underlying dynamics of most ecological processes are unknown and instead
must be inferred from this (sparse) data. Much of the history of
quantitatively modeling ecosystems have been done in the language of
dynamical systems, describing how the value of an observable state of
the system, represented by a vector of numbers
\([x_1, x_2, \dots, x_n]^T = \vec{x}\) changes as over time, yielding
models in the form of differential equations in continuous-time
settings, \(\frac{dx}{dt} = f(x)\), or difference equations in
discrete-time settings, \(x_t = f(x_{t-1})\), where
\(f:\mathbb{R}^n \to \mathbb{R}^n\) is an arbitrary function describing
how the system changes on a moment-to-moment basis (e.g.~in the context
of communities, \(f\) could be Lotka-Voltera, Holling-Type-III or
DeAngelis-Beddington functional response). The form of this functional
response in real systems, and whether it is meaningfully non-zero for a
given species interaction, is effectively unknown, and some forms are
inherently more ``forecastable'' than others (Beckage \emph{et al.}
2011; Chen \emph{et al.} 2019; Pennekamp \emph{et al.} 2019). The
initial success of these forms of models can be traced back to the
larger program of ontological reductionism, which became the default
approach to modeling in the sciences after its early success in physics,
which, by the time ecology was becoming a quantitative science (sometime
in the 20th century, depending on who you ask), became the foundation
for mathematical models in ecology.

@@

\begin{figure}
\hypertarget{fig:foodwebtraj}{%
\centering
\includegraphics{./figures/foodwebtraj.png}
\caption{Sample output of simulated food web dynamics from
MetacommunityDynamics.jl}\label{fig:foodwebtraj}
}
\end{figure}

@@

Biodiversity science would be wise to use the success of
numerical-weather prediction and the Earth monitoring system that
supports it as a template. Further, we should embrace methodological
advances in computational statistics that can enable more robust
prediction of the dynamics of complex systems than is possible using the
tools of analytic models---not just because they can provide more
efficacious predictions, but also because they enable a more direct
representation of uncertainty, which is crucial for developing
ecological forecasting models which provide actionable information to
stakeholders.

>>>>

The second major challenge in ecological forecasting is that the
underlying dynamics of most ecological processes are unknown and instead
must be inferred from this (sparse) data. Much of the history of
quantitatively modeling ecosystems have been done in the language of
dynamical systems, describing how the value of an observable state of
the system, represented by a vector of numbers
\([x_1, x_2, \dots, x_n]^T = \vec{x}\) changes as over time, yielding
models in the form of differential equations in continuous-time
settings, \(\frac{dx}{dt} = f(x)\), or difference equations in
discrete-time settings, \(x_t = f(x_{t-1})\), where
\(f:\mathbb{R}^n \to \mathbb{R}^n\) is an arbitrary function describing
how the system changes on a moment-to-moment basis (e.g.~in the context
of communities, \(f\) could be Lotka-Voltera, Holling-Type-III or
DeAngelis-Beddington functional response). The form of this functional
response in real systems, and whether it is meaningfully non-zero for a
given species interaction, is effectively unknown and must be predicted
(\textbf{Strydom2021RoaMap?}), and some forms are inherently more
``forecastable'' than others (Beckage \emph{et al.} 2011; Chen \emph{et
al.} 2019; Pennekamp \emph{et al.} 2019). The initial success of these
forms of models can be traced back to the larger program of ontological
reductionism, which became the default approach to modeling in the
sciences after its early success in physics, which, by the time ecology
was becoming a quantitative science (sometime in the 20th century,
depending on who you ask), became the foundation for mathematical models
in ecology.

@@

\begin{figure}
\hypertarget{fig:foodwebtraj}{%
\centering
\includegraphics{./figures/foodwebtraj.png}
\caption{Sample output of simulated food web dynamics from
MetacommunityDynamics.jl. Timestep (x-axis), and biomass of each species
(y-axis).}\label{fig:foodwebtraj}
}
\end{figure}

@@

Biodiversity science would be wise to use the success of
numerical-weather prediction and the Earth monitoring system that
supports it as a template. Further, we should embrace methodological
advances in computational statistics that can enable more robust
prediction of the dynamics of complex systems than is possible using the
tools of analytic models---not just because they can provide more
efficacious predictions, but also because they enable a more direct
representation of uncertainty, which is crucial for developing
ecological forecasting models which provide actionable information to
stakeholders, and to find a way for humans to live sustainably on Earth,
the only planet on which life has ever been known to occur.
